The Weakly Nonlinear Magnetorotational Instability

### 2016-09-02 paper/widegap_mri_paper.tex

\section{Discussion}
\label{sec:discussion}
Our work should be placed in the broader context of emergent pattern formation in physical systems. The real Ginzburg-Landau equation derived here governs the slow-parameter evolution of the standard MRI close to threshold. The GLE arises in a number of other physical systems, and in each case it is a consequence not of the particular physics at hand, but of the underlying symmetries in the problem. Here we make a phenomenological comparison to two other systems that give rise to a GLE. The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}). If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric. %When the system eventually becomes unstable to convective heat transfer it undergoes a symmetry breaking, allowing the 
At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance. Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the TC device as an infinitely long cylinder. The MRI breaks this symmetry, forming cells along the vertical length of the domain. Just as Rayleigh-B\'enard cells transport heat vertically, the MRI cells transport angular momentum horizontally. The symmetry breaking of each of these systems is described near onset by the real GLE.

>>>>

\section{Discussion}
\label{sec:discussion}
In this work we carry out a formal weakly nonlinear multiscale analysis of the MRI in a Taylor Couette flow. We analyze both the standard and helical MRI, which differ only in the geometry of their imposed background magnetic fields. We find that the amplitude function, which governs the behavior of the system on long length- and timescales, obeys a real GLE for the case of the standard MRI, and a complex GLE for the helical MRI. These two systems are thus subject to different large-scale phase dynamics.

Our work should be placed in the broader context of emergent pattern formation in physical systems. The real Ginzburg-Landau equation derived here governs the slow-parameter evolution of the standard MRI close to threshold. The GLE arises in a number of other physical systems, and in each case it is a consequence not of the particular physics at hand, but of the underlying symmetries in the problem. Here we make a phenomenological comparison to two other systems that give rise to a GLE. The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}). If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric. %When the system eventually becomes unstable to convective heat transfer it undergoes a symmetry breaking, allowing the 
At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance. Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the TC device as an infinitely long cylinder. The MRI breaks this symmetry, forming cells along the vertical length of the domain. Just as Rayleigh-B\'enard cells transport heat vertically, the MRI cells transport angular momentum horizontally. The symmetry breaking of each of these systems is described near onset by the real GLE.
